---
title: "FinalProject"
author: "Fjolle Gjonbalaj"
date: "12/04/2021"
output: html_document
---


#Introduction

The perception of corruption in a country is a crucial component when it comes to the happiness of the people and their healthy life expectancy. However, corruption is not the only indicator of happiness. Other variables such as social support, freedom to make life choices, generosity as well as the log GDP per capita also play a key role on the happiness of a nation. 


Sources: https://worldhappiness.report/


```{r , echo = FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
happiness <- read.csv("https://raw.githubusercontent.com/Fjolle/Data-Mining-Project/main/Data/world-happiness-report-2021.csv")
happiness <- na.omit(world_happiness_report_2021)
library(tidyverse)
```


In order to construct an accurate statictical model that accurately predicts happiness levels across
the world I apply four different Healthy_life_expectancy statistical models: The linear model, Lasso, Random Forest and Boosting.
Predictions using each one of the above models are then compared to observe which one provides us with
the best performance. I analyze different Healthy_life_expectancy broken-down regions of the world in order to observe the scoreof the factors adding to the population's happiness level.

#Healthy life expectancy explained by perceptions of corruption

```{r, echo = FALSE, warning=FALSE}
happiness <- read.csv("https://raw.githubusercontent.com/Fjolle/Data-Mining-Project/main/Data/world-happiness-report-2021.csv")
happiness %>%
  group_by(Regional_indicator) %>%
  ggplot() +
  geom_line(aes(y = Healthy_life_expectancy, x = Explained_by_Perceptions_of_corruption,
  color = Regional_indicator)) +
  facet_wrap(~ Regional_indicator, scales = "free") + 
  labs(x = "Regional Indicator", y = "Average Life Expectancy",
       title = "Figure 1: Healthy Life Expectancy based on Freedom to make life choices",
       caption = "Data from Gallup World Poll")+
  theme_minimal()
```


#Healthy life expectancy explained by social support
```{r, echo = FALSE, warning=FALSE}
happiness <- read.csv("https://raw.githubusercontent.com/Fjolle/Data-Mining-Project/main/Data/world-happiness-report-2021.csv")
happiness %>%
  group_by(Regional_indicator) %>%
  ggplot() +
  geom_line(aes(y = Healthy_life_expectancy, x = Social_support, color = Regional_indicator)) +
  facet_wrap(~ Regional_indicator, scales = "free") + 
  labs(x = "Regional Indicator", y = "Average Life Expectancy",
       title = "Figure 2:Healthy Life Expectancy based on Freedom to make life choices",
       caption = "Data from Gallup World Poll")+
  theme_minimal()
```


#Healthy life expectancy explained by freedom to make life choices
```{r, echo = FALSE, warning=FALSE}
happiness <- read.csv("https://raw.githubusercontent.com/Fjolle/Data-Mining-Project/main/Data/world-happiness-report-2021.csv")
happiness %>%
  group_by(Regional_indicator) %>%
  ggplot() +
  geom_line(aes(y = Healthy_life_expectancy, x = Explained_by_Freedom_to_make_life_choices, color = Regional_indicator)) +
  facet_wrap(~ Regional_indicator, scales = "free") + 
  labs(x = "Regional Indicator", y = "Average Life Expectancy",
       title = "Figure 3: Healthy Life Expectancy based on Freedom to make life choices",
       caption = "Data from Gallup World Poll")+
  theme_minimal()
```


#Healthy life expectancy explained by generosity
```{r, echo = FALSE, warning=FALSE}
happiness <- read.csv("https://raw.githubusercontent.com/Fjolle/Data-Mining-Project/main/Data/world-happiness-report-2021.csv")
happiness %>%
  group_by(Regional_indicator) %>%
  ggplot() +
  geom_line(aes(y = Healthy_life_expectancy, x = Explained_by_Generosity, color = Regional_indicator)) +
  facet_wrap(~ Regional_indicator, scales = "free") + 
  labs(x = "Regional Indicator", y = "Average Life Expectancy",
       title = "Figure 4: Healthy Life Expectancy based on Freedom to make life choices",
       caption = "Data from Gallup World Poll")+
  theme_minimal()
```


#Healthy life expectancy explained by log GDP per capita
```{r, echo = FALSE, warning=FALSE}
happiness <- read.csv("https://raw.githubusercontent.com/Fjolle/Data-Mining-Project/main/Data/world-happiness-report-2021.csv")
happiness %>%
  group_by(Regional_indicator) %>%
  ggplot() +
  geom_line(aes(y = Healthy_life_expectancy, x = Explained_by_Log_GDP_per_capita, color = Regional_indicator)) +
  facet_wrap(~ Regional_indicator, scales = "free") + 
  labs(x = "Regional Indicator", y = "Average Life Expectancy",
       title = "Figure 5: Healthy Life Expectancy based on Freedom to make life choices",
       caption = "Data from Gallup World Poll")+
  theme_minimal()
```


# First Model: Hand-Built Linear Model
```{r, echo = FALSE, warning=FALSE}
happiness <- read.csv("https://raw.githubusercontent.com/Fjolle/Data-Mining-Project/main/Data/world-happiness-report-2021.csv")
lm_happiness = lm(Healthy_life_expectancy ~ Regional_indicator + Explaine_by_Social_support+
                    Explained_by_Log_GDP_per_capita+Explained_by_Freedom_to_make_life_choices+ 
                    Explained_by_Generosity+Explained_by_Perceptions_of_corruption+
                    Regional_indicator*Explaine_by_Social_support+
                    Regional_indicator*Explained_by_Log_GDP_per_capita+
                    Regional_indicator*Explained_by_Freedom_to_make_life_choices+
                    Regional_indicator*Explained_by_Generosity+
                    Regional_indicator*Explained_by_Perceptions_of_corruption,data=happiness)

```


#Second Model: Lasso 
```{r, echo = FALSE, warning=FALSE}
happiness <- read.csv("https://raw.githubusercontent.com/Fjolle/Data-Mining-Project/main/Data/world-happiness-report-2021.csv")
library(glmnet)

x = model.matrix(Healthy_life_expectancy~Regional_indicator+Explaine_by_Social_support+
                   Explained_by_Log_GDP_per_capita+Explained_by_Freedom_to_make_life_choices+ 
                   Explained_by_Generosity+Explained_by_Perceptions_of_corruption, data=happiness)[,-1]
x = scale(x, center=TRUE, scale=TRUE) 
y = happiness$Healthy_life_expectancy
grid=15^seq(5,-1, length =30)
lasso.mod=glmnet(x,y,alpha=1, lambda =grid)
cv.out=cv.glmnet(x,y,alpha=1)
bestlam =cv.out$lambda.min
plot(lasso.mod)
title("Lasso Coefficients", line = 3)

library(data.table)
library(kableExtra)
lasso.coef=predict(lasso.mod ,type ="coefficients",s=bestlam)
LassoCoef=as.data.table(as.matrix(lasso.coef), keep.rownames = TRUE)
kable(LassoCoef, col.names = c("Predictor", "Estimate"), caption = "Lasso Model Predictor Estimates",  format_caption = c("italic", "underline")) %>%
  kable_styling(bootstrap_options = "striped", full_width = F)

#Split into training and testing sets
library(rsample)
happiness_split = initial_split(happiness, prop = 0.8)
happiness_train = training(happiness_split)
happiness_test = testing(happiness_split)
#K- fold cross validation
library(modelr)
happiness_folds = crossv_kfold(happiness, k=20)
best = map(happiness_folds$train, ~ lm(Healthy_life_expectancy~Regional_indicator+
                                         Explaine_by_Social_support+
                                         Explained_by_Log_GDP_per_capita+
                                         Explained_by_Freedom_to_make_life_choices+ 
                                         Explained_by_Generosity+
                                         Explained_by_Perceptions_of_corruption,
                                       data=happiness_train))


library(mosaic)
# Mean RMSE
rmse = function(y, yhat) {
  sqrt(mean((y - yhat)^2))}
LoopModels = do(20)*{
  n = nrow(happiness)
  n_train = round(0.8*n)  
  n_test = n - n_train
  train_cases = sample.int(n, n_train, replace=FALSE)
  test_cases = setdiff(1:n, train_cases) 
  happiness_train = happiness[train_cases,]
  happiness_test = happiness[test_cases,]
  
  lm_happiness = lm(Healthy_life_expectancy~Regional_indicator+
                      Explaine_by_Social_support+
                      Explained_by_Log_GDP_per_capita+
                      Explained_by_Freedom_to_make_life_choices+ 
                      Explained_by_Generosity+
                      Explained_by_Perceptions_of_corruption
                    , data=happiness_train)
  
  x = model.matrix(Healthy_life_expectancy~Regional_indicator+Explaine_by_Social_support+
                     Explained_by_Log_GDP_per_capita+Explained_by_Freedom_to_make_life_choices+ 
                     Explained_by_Generosity+Explained_by_Perceptions_of_corruption, data=happiness)[,-1]
  
  x = scale(x, center=TRUE, scale=TRUE) 
  y = happiness$Healthy_life_expectancy
  grid=10^seq(10,-2, length =100)
  lasso.mod=glmnet(x[train_cases,],y[train_cases],alpha=1, lambda =grid)
  cv.out=cv.glmnet(x[train_cases,],y[train_cases],alpha=1)
  bestlam =cv.out$lambda.min
  
  yhat_test_lmhappiness = predict(lm_happiness, happiness_test)
  yhat_Lasso = predict(lasso.mod, s=bestlam, newx=x[test_cases,])
  
  c(RMSEhappiness = rmse(happiness_test$Healthy_life_expectancy, yhat_test_lmhappiness),
    RMSELasso = rmse(happiness_test$Healthy_life_expectancy,yhat_Lasso))
  
}
RMSEMeans = c("Hand-Built Linear Model" = mean(LoopModels$RMSEhappiness), 
              "Lasso" = mean(LoopModels$RMSELasso))
kable(RMSEMeans, col.names = c("Mean RMSE"), caption = "Mean RMSE",  format_caption = c("italic", "underline")) %>%
  kable_styling(bootstrap_options = "striped", full_width = F)

```


##Bagging
```{r, echo = FALSE, warning=FALSE}
happiness <- read.csv("https://raw.githubusercontent.com/Fjolle/Data-Mining-Project/main/Data/world-happiness-report-2021.csv")
library(randomForest)
set.seed(1)
happyBagging = randomForest(Healthy_life_expectancy~Regional_indicator+Explaine_by_Social_support+
                              Explained_by_Log_GDP_per_capita+Explained_by_Freedom_to_make_life_choices+ 
                              Explained_by_Generosity+Explained_by_Perceptions_of_corruption, data = happiness_train, mtry=7, importance=TRUE)
happyBagging


yhat_happyBagging = predict(happyBagging, newdata = happiness_test)
plot(yhat_happyBagging, happiness_test$Healthy_life_expectancy, xlab = "Predicted Values for Healthy life expectancy: Bagging", ylab = "Healthy_life_expectancy")
title("Comparison between Bagging Predicted Healthy life expectancy and Actual Healthy life expectancy")
```


#Random Forest
```{r, echo = FALSE, warning=FALSE}
happiness <- read.csv("https://raw.githubusercontent.com/Fjolle/Data-Mining-Project/main/Data/world-happiness-report-2021.csv")
set.seed(1)
happyRandomForest = randomForest(Healthy_life_expectancy~Regional_indicator+Explaine_by_Social_support+
                                   Explained_by_Log_GDP_per_capita+Explained_by_Freedom_to_make_life_choices+ 
                                   Explained_by_Generosity+Explained_by_Perceptions_of_corruption, data = happiness_train,
                                 mtry=3, importance=TRUE)
happyRandomForest = predict(happyRandomForest, newdata = happiness_test)
plot(happyRandomForest, happiness_test$Healthy_life_expectancy, xlab = "Predicted Values for Healthy life expectancy: Random Forest", ylab = "Healthy life expectancy")
title("Comparison between Random Forest Predicted Healthy life expectancy and Actual Healthy life expectancy")
```




### Comparison of all 4 Predictive Models: Hand-Built Linear Model, Lasso, Bagging and Random Forest:
```{r, echo = FALSE, warning=FALSE}
happiness <- read.csv("https://raw.githubusercontent.com/Fjolle/Data-Mining-Project/main/Data/world-happiness-report-2021.csv")
library(foreach)
N=nrow(happiness)
K=3
fold_id = rep_len(1:K, N)
fold_id = sample(fold_id, replace = FALSE)

#Compute cross validation error for Model1 
err_save = rep(0, K)
for (i in 1:K) {
  train_set = which(fold_id != i)
  y_testCV = happiness$Healthy_life_expectancy[-train_set]
  lm_Healthy_life_expectancyCV = lm(Healthy_life_expectancy ~  Regional_indicator + Explaine_by_Social_support+
                                      Explained_by_Log_GDP_per_capita+Explained_by_Freedom_to_make_life_choices+ 
                                      Explained_by_Generosity+Explained_by_Perceptions_of_corruption+
                                      Regional_indicator*Explaine_by_Social_support+
                                      Regional_indicator*Explained_by_Log_GDP_per_capita+
                                      Regional_indicator*Explained_by_Freedom_to_make_life_choices+
                                      Regional_indicator*Explained_by_Generosity+
                                      Regional_indicator*Explained_by_Perceptions_of_corruption 
                                    , data=happiness[train_set,]) 
  
  yhat_Healthy_life_expectancy_CV = predict(lm_Healthy_life_expectancyCV, newdata = happiness[-train_set,])
  
  err_save[i] = mean((y_testCV - yhat_Healthy_life_expectancy_CV)^2)
}
RMSE = sqrt(mean(err_save))
RMSE


#compute cross validation error for Lasso Model 
err_saveLasso = rep(0, K)
for (i in 1:K) {
  train_set = which(fold_id != i)
  y_testCV = happiness$Healthy_life_expectancy[-train_set] 
  x = model.matrix(Healthy_life_expectancy~Regional_indicator+Explaine_by_Social_support+
                     Explained_by_Log_GDP_per_capita+Explained_by_Freedom_to_make_life_choices+ 
                     Explained_by_Generosity+Explained_by_Perceptions_of_corruption, data=happiness)[,-1]
  x = scale(x, center=TRUE, scale=TRUE) 
  y = happiness$Healthy_life_expectancy
  grid=10^seq(10,-2, length =100)
  lasso.mod=glmnet(x[train_set,],y[train_set],alpha=1, lambda =grid)
  cv.out=cv.glmnet(x[train_set,],y[train_set],alpha=1)
  bestlam =cv.out$lambda.min
  
  yhat_lm_Healthy_life_expectancy_Lasso = predict(lasso.mod, s=bestlam, newx=x[-train_set,])
  
  err_saveLasso[i] = mean((y_testCV - yhat_lm_Healthy_life_expectancy_Lasso)^2)
}
RMSE3 = sqrt(mean(err_saveLasso))
RMSE3

#Bagging Model
err_saveTreeBag = rep(0, K)
for (i in 1:K) {
  train_set = which(fold_id != i)
  happiness_train = happiness[train_set,]
  #train_set = scale(train_set, center=TRUE, scale=TRUE) 
  y_testCV = happiness$Healthy_life_expectancy[-train_set] 
  y = happiness$Healthy_life_expectancy
  happiness_test = happiness[-train_set,]
  
  set.seed(1)
  greenbuildBagging = randomForest(Healthy_life_expectancy~Regional_indicator+Explaine_by_Social_support+
                                     Explained_by_Log_GDP_per_capita+Explained_by_Freedom_to_make_life_choices+ 
                                     Explained_by_Generosity+Explained_by_Perceptions_of_corruption, data = happiness_train,
                                   mtry=17, importance=TRUE)
  
  yhat_greenbuildBagging = predict(greenbuildBagging, newdata = happiness_test)
  
  err_saveTreeBag[i] = mean((y_testCV - yhat_greenbuildBagging)^2)
}
RMSE4 = sqrt(mean(err_saveTreeBag))
RMSE4 


#Random Forest
err_saveTreeForest = rep(0, K)
for (i in 1:K) {
  train_set = which(fold_id != i)
  happiness_train = happiness[train_set,]
  #train_set = scale(train_set, center=TRUE, scale=TRUE) 
  y_testCV = happiness$Healthy_life_expectancy[-train_set] 
  y = happiness$Healthy_life_expectancy
  happiness_test = happiness[-train_set,]
  
  set.seed(1)
  greenbuildRandomForest = randomForest(Healthy_life_expectancy~Regional_indicator+Explaine_by_Social_support+
                                          Explained_by_Log_GDP_per_capita+Explained_by_Freedom_to_make_life_choices+ 
                                          Explained_by_Generosity+Explained_by_Perceptions_of_corruption, data = happiness_train,
                                        mtry=4, importance=TRUE)
  
  yhat_greenbuildRandomForest = predict(greenbuildRandomForest, newdata = happiness_test)
  
  err_saveTreeForest[i] = mean((y_testCV - yhat_greenbuildRandomForest)^2)
}
RMSE5 = sqrt(mean(err_saveTreeForest))
RMSE5

AvgRMSEModels = c("LOOCV RMSE Healthy_life_expectancy Hand-Built Model"=sqrt(mean(err_save)), 
                  "LOOCV RMSE Model Lasso Model" = sqrt(mean(err_saveLasso)),
                  "LOOCV RMSE Model Bagging Model" = sqrt(mean(err_saveTreeBag)),
                  "LOOCV RMSE Model RandomForest Model" = sqrt(mean(err_saveTreeForest)))
kable(AvgRMSEModels, col.names = c("LOOCV RMSE"), caption = " LOOCV RMSE per Model",  format_caption = c("italic", "underline")) %>%
  kable_styling(bootstrap_options = "striped", full_width = F)


set.seed(1)
greenbuildRandomForest2 = randomForest(Healthy_life_expectancy~Regional_indicator+Explaine_by_Social_support+
                                         Explained_by_Log_GDP_per_capita+Explained_by_Freedom_to_make_life_choices+ 
                                         Explained_by_Generosity+Explained_by_Perceptions_of_corruption, data = happiness,
                                       mtry=4, importance=TRUE)
VariableImp = as.data.table(importance(greenbuildRandomForest2), keep.rownames = TRUE)
kable(VariableImp, caption = "**Table 1.4 Variable Importance in Random Forest Model**", col.names = c("Predictor", "% Increase in MSE", "Increase in Node Purity"),  format_caption = c("italic", "underline")) %>%
  kable_styling(bootstrap_options = "striped", full_width = F)
```